---
title: "R Notebook"
output: html_notebook
---

# 1. What is vector?

-Vectors are 2 types: atomic, list.

-Every vectors has 2 features: typeof, length

-There are 4 increased vectors: Factors, datetime, dataframe, tibbles

## Important Atomic Vectors

```{r}

c(TRUE, FALSE, NA)

# Integer & Numeric

a<- 1

b<- 1L

# Characters

c<- "John"

```


# 2. Vector Formating

## Creating Vectors

```{r}

a<- c(1,2,3,4)

b<- c("a","b","c")

c<- 1:10

d<- 1

```

## Combining Vectors

```{r}

v<- 1:10

y<- 2:11

v+y

z<- 1:5

v+z

c(v,y)

c(v,z)

c(v,y,z)

rbind (v,y) # r stands for row

cbind (v,y) # c stands for column

```

## Vector Naming

```{r}

n<- c(x = 1, y = 2, z = 3)

n
```

## Vector Types and Converting

### Vector Types

```{r}

is.logical(a)
is.integer(a)
is.double(a)
is.numeric()
is.character()
is.atomic()
is.list()
is.vector()


```

### Converting Vector Types

```{r}

tf<- c("T", "F")

tf<- as.logical(tf)

p<- 1

typeof(p)

p<- as.character(p)

typeof(p)

p

r<- c(1,2,"a")
r<- as.character(r)
r

s<- c(10,10,10,20,20,30,40,50,50,50)
s<- factor(s)
s
# Let's convert s back to vector again!

as.numeric(as.character(s))
```


# 3. Vector Element Operations

## Vector Index and Element selection

```{r}

vektor<- rnorm(30) # it generates 30 numbers by normal distribution btw -3 and 3

vektor[1]
vektor[c(2,3,4)]
vektor[5:15]

```

## Assign Value using Index

```{r}

vektor[1]<- 1
vektor[1]
```

## Assign Collective Value

```{r}

vektor[c(2,3,4)]<- c(3,5,7)
vektor[c(2,3,4)]

vektor[5:10]<- 0
vektor[5:10]



```

##Record Selection from Named Vector

```{r}

vv<- c(x=1, y=2, z=3)

vv["x"]
vv[c("x","z")]

```

## Conditional Record Selection

```{r}

vektor
vektor[vektor>=0]

vektor[vektor>0 & vektor<1]

```

## Erasing Records of Vector

```{r}

yy<- 1:10

yy<- NULL

# The other method is to slice

yy<- 1:10

yy<-yy[2:6]
yy

```


# 4. Vectoral Calculations

## Mathematical Operations

```{r}

v<- 1:10

y<- 2:11

v+y

v-y

v*y

v/y

sqrt(v)
sum(v)
mean(v)
median(v)


v^2

```

## Statistical Operations

```{r}

min(v)
max(v)
mean(v)
median(v)
cumsum(v)


```

