---
title: "7.Data Manipulation"
author: "Vedat Ozkan"
date: "December 3, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


select()   : variable selection
filter()   : observation selection
mutate()   : varibale creation
summarise(): summary statistics
group_by() : grouping
left_join(): join operration
summarise(): data summary


# 0. Data Set

nycflights13: flight data of NYC in 2013

https://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=120&Link=0

Understanding dataset

```{r}

library(dplyr)
library(nycflights13)

df<- flights

dim(df)

head(df)

df

str(df)

summary(df)

```


```{r}
# Let's find the number of flight and average delay time by carrier (flight companies). Ignore less than 10 minutes

df %>% 
  select(dep_delay, day, carrier) %>%
  filter(dep_delay > 10) %>%
  group_by(carrier) %>%
  summarise(n = n(),avg = mean(dep_delay) )

```


# 1. Variable Selection Operations: select()

select()
```{r}

# in order to select

select(df, carrier, flight, tailnum)

select(df, carrier, origin:hour)

select(df, origin:hour)

select(df, 1:3)

# in order to remove

select(df, -carrier)

select(df, -c(carrier, tailnum))

select(df, -c(origin:hour))


```


## contains()

```{r}

select(df, contains("time"))


```


## starts_with()

```{r}

select(df, starts_with("dep"))

```


## ends_with()

```{r}

select(df, ends_with("delay"))

```



## num_range()

```{r}

m<- matrix(1:25, 5,5)

colnames(m)<- paste("x",1:5, sep = "")

#IMPORTANT: First argument of Dplyr's functions has to be Dataframe !!

select(data.frame(m), num_range("x",2:4))

```


Also, sub-functions of select can be used in a combo way

```{r}

select(df,carrier, tailnum, contains("time"))


```



# 2.Observation Operations: filter()

filter()
```{r}

filter(df, year == 2013 & month == 2)

filter(df, month ==2 & day ==18 )

filter(df,dep_delay > 1000)

filter(df, dep_delay > mean(df$dep_delay, na.rm = TRUE))

filter(df, dep_delay > mean(df$dep_delay, na.rm = TRUE) + sd(df$dep_delay, na.rm = TRUE) )

```


## arrange()

```{r}


a<- select(df, dep_delay, month, day)

arrange(a, dep_delay, month, day) # ascending sort

#-----

arrange(a, desc(dep_delay), month, day) # descending sort

```


```{r}

df %>%
  filter(month == 2, day == 18) %>%
  select(dep_delay, month) %>%
  arrange(desc(dep_delay), month)

```



## sample_frac() 

This is for sample choosing randomly by ratio
```{r}

df %>% sample_frac(0.05)

```


## sample_n 

This is for sample choosing randomly by number
```{r}

df %>% sample_n(100)

```



## slice()

```{r}

df %>% slice(50:100)

```



## top_n()

```{r}

df %>% top_n(10)

```



# 3. Variable Transformation: mutate()

mutate()
```{r}

# Suppose we want to search the flights depart delayed, but arrive earlier, also speed

sample<- df %>%
            sample_n(1000) %>%
            select(arr_delay, dep_delay, distance, arr_time)


mutate(sample,
       difference = arr_delay - dep_delay,
       speed = distance / arr_time * 60,
       new = difference / speed)


```



## rename()

```{r}
# in order to rename variables

rename(df, 
       arrival_delay = arr_delay,
       departure_delay = departure_delay
       )


```



## transmute()

```{r}

# It provides to add new variables whereas current variables remove

transmute(sample,
       difference = arr_delay - dep_delay,
       speed = distance / arr_time * 60,
       new = difference / speed)


```



# 4. Grouping: group_by()

```{r}

# summarise() is used with group_by together

df %>%
  group_by(carrier) %>%
  summarise(num = n(),
            avg = mean(dep_delay, na.rm = TRUE),
            med = median(dep_delay, na.rm = TRUE),
            std = sd(dep_delay, na.rm = TRUE),
            var = var(dep_delay, na.rm = TRUE),
            avg_distance = mean(distance)
            )

```



# 5. Data Summary : summarise()


## counts

## mean(), median(), sum(), sd(), var()

## first(), last(), nth()

## quantile(), min(), max()




# 6. Join Operations


## inner_join()

## left_join()

## right_join()

## full_join



# 7. Tidy Data

```{r}

library(tidyverse)

table1

table2

table3

table4a

table4b

table5

```



## gathering and spreading



## gathering


```{r}

table4a %>% 
  gather('1999',"2000", key = "year", value = 'cases')


```


```{r}

table4b %>%
  gather("1999","2000", key = "year", value = "population")

```


```{r}

# Let's merge table4a with table4b

tidy_a<- table4a %>% 
  gather('1999',"2000", key = "year", value = 'cases')


tidy_b<- table4b %>%
  gather("1999","2000", key = "year", value = "population")


left_join(tidy_a, tidy_b)


```



## spreading 

spread() is opposite of gather()

```{r}

table2

spread(table2, key = "type", value = "count")

```




## separating and uniting


## separate()

```{r}

table3 %>%
  separate(rate, into = c("cases","population"), sep ="/", convert = TRUE)


```



## unite()

```{r}

table5 %>%
  unite(new, century, year, sep = "")

```





