---
title: "Visualization_ggplot2.Rmd"
author: "Vedat Ozkan"
date: "December 19, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Resources:

http://r4ds.had.co.nz
https://github.com/jtleek


```{r}

# Overview


library(tibble)
library(dplyr)
library(ggplot2)


df<- tibble(
  use = factor(c("wd","we"),levels = c("wd","we")),
  duration = c(100,180)
)


a<- ggplot(df, aes(x = use, y = duration)) +
  geom_bar(stat = "identity")

a+geom_bar(stat = "identity")


a + geom_bar(colour = "black",
             fill = "orange",
             stat = "identity"
           ) +
  guides(fill = FALSE) +
  xlab("Wekdays - Weekend") +
  ylab("Usage Duration") +
  ggtitle("Phone Usage Statistics")
  
# Note: Below provides to see the last error

library(rlang)
last_error()

```


# 1. Univariant Visualization


## 1.1. Discrete Variable

In order to visualize Discrete Variable is needed to visualize, you should do it by:

a) by frequency
b) by its absolute value

```{r}

df<- tibble(
  use = c(rep("wd", 20), rep("we", 10))
)

df1<- tibble(
  use = factor(c("wd","we"),levels = c("wd","we")),
  duration = c(100,180)
)

```


### Bar Graph: geom_bar()

```{r}

# for df

ggplot(df,aes(use)) +
  geom_bar()

# for df1

ggplot(df1, aes(x = use, y = duration)) +
  geom_bar(stat = "identity")


```



### Line Graph: geom_line()

```{r}

df <- tibble(
  use = factor(c("wd","we"),levels = c("wd","we")),
  duration = c(100,180)
)


ggplot(df, aes(x=use, y = duration, group = 1))+
  geom_line() 

# let's add geom_point() function

ggplot(df, aes(x=use, y = duration, group = 1))+
  geom_line() +
  geom_point()

# let's add a few more feature

ggplot(df, aes(x=use, y = duration, group = 1))+
  geom_line(color = "red", linetype = "dashed", shape =1.6) +
  geom_point(color = "red", size = 4, shape = 21, fill = "blue")+
  expand_limits(y = 0)

```



### Pie Graph

```{r}








```




## 1.2. Continuous Variable


```{r}

df<- tibble(
  
  gender = factor(rep(c("M","F"), each = 500)),
  duration = c(rnorm(500, mean = 80), rnorm(500, mean = 81))
)

df

```


### Histogram: geom_histogram

```{r}

ggplot(df, aes(duration)) +
  geom_histogram()



ggplot(df, aes(duration)) +
  geom_histogram(binwidth = .5, 
                 color ="black", 
                 fill = "orange") +
  xlab("Duration Distribution") +
  ylab("Frequency") + 
  ggtitle("Histogram")


```


Density Graph: geom_density()

```{r}

  ggplot(df, aes(duration)) +
  geom_density()
  
# ----------
  
  ggplot(df, aes(duration)) +
  geom_histogram(aes(y = ..density..), 
                 binwidth = .5, 
                 color ="black",
                 fill = "orange") +
  geom_density()

# -------------

ggplot(df, aes(duration)) +
  geom_histogram(aes(y = ..density..), 
                 binwidth = .5, 
                 color ="black",
                 fill = "orange") +
  geom_density(alpha = .3,
               fill = "black")


```


```{r}

# How to add GENDER dimension to Histogram 

ggplot(df, aes(duration, fill = gender)) +
  geom_histogram(binwidth = .5, 
                 alpha = .5,
                 position = "identity"
                )


# How to add GENDER dimension to Density

ggplot(df, aes(duration, fill = gender)) +
  geom_density( binwidth = .5, 
                alpha = .5,
                position = "identity"
                
                )


```


facet_grid()

It provides to divide data into 2 more dimensions
```{r}
# How to display dimensions in 2 different graph


ggplot(df, aes(duration)) +
  geom_histogram(binwidth = .5, 
                 color = "black",
                 fill = "orange") +
  facet_grid(gender ~ .)

# facet_grid() above ensures to display 2 histogram by male and female separately

```



### Box Plot: geom_boxplot()

```{r}

ggplot(df, aes(x = gender, y = duration)) +
  geom_boxplot()

# Let's fill the baxes with  different color

ggplot(df, aes(x = gender, y = duration, fill = gender)) +
  geom_boxplot()

# Let's remove autometed explanation in visualization

ggplot(df, aes(x = gender, y = duration, fill = gender)) +
  geom_boxplot() +
  guides(fill = FALSE)


# Let's change the direction of axis

ggplot(df, aes(x = gender, y = duration, fill = gender)) +
  geom_boxplot() +
  guides(fill = FALSE) +
  coord_flip()


```



### Area Graph: geom_area()

```{r}

ggplot(df, aes(duration))+
  geom_area(stat = "bin")

```



### Frequency Graph: geom_preqpoly()

```{r}

 ggplot(df, aes(duration)) +
  geom_freqpoly()

```



### Dotplot Graph: geom_dotplot()

```{r}

ggplot(df, aes(duration)) +
  geom_dotplot(binwidth = 0.1)

```


# 2.Bivariate Visualization


## 2.1. Continuous X, Continuous Y



```{r}
# The amount of time women and men spend on the internet and mobile

df<- tibble(
  gender = factor(rep(c("M","F"), each = 500)),
  mobile = c(rnorm(500, mean = 80), rnorm(500, mean = 82)),
  web    = c(rnorm(500, mean = 82), rnorm(500, mean = 81))
)

df

```



### Scatter-plot

```{r}

ggplot(df,aes(mobile, web)) +
  geom_point()

# -------- let's add linear regression line ---

ggplot(df,aes(mobile, web)) +
  geom_point() +
  geom_smooth(method = lm)

# -------let's remove confidence interval in visualization---

ggplot(df, aes(mobile, web)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

# ----- if we want to see non-linear regression line ---

ggplot(df, aes(mobile, web)) +
  geom_point() +
  geom_smooth()

# ----- let's do another thing, let's display Male and Female observations in the graph ----

ggplot(df,aes(mobile, web, color = gender)) +
  geom_point()

# ---- Now, let's think of linear regression lines for  Male and female seperately ---

ggplot(df,aes(mobile, web, color = gender)) +
  geom_point() +
  geom_smooth(method = lm)

```


## 2.2. Discrete X, COntinuous Y

```{r}

df<- tibble(
  use = factor(c("wd","we"), levels = c("wd","we")),
  duration = c(100, 180)
)

df

```



### Bar Graph: geom_bar()


```{r}

ggplot(df,aes(use, duration, fill = use)) +
  geom_bar(color =  "black", 
           fill = "orange", 
           stat = "identity")

```



### Box Plot: geom_boxplot()

```{r}


df<- tibble(
  
  gender = factor(rep(c("M","F"), each = 500)),
  duration = c(rnorm(500, mean = 80), rnorm(500, mean = 81))
)


ggplot(df, aes(x=gender, y=duration, fill = gender))+
  geom_boxplot()

```



### Dot Plot: geom_dotplot()

```{r}

ggplot(df, aes(x=gender, y=duration, fill = gender))+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               binwidth = 0.1
               )

```



### Violin Graph: geom_violin()

```{r}

ggplot(df, aes(x=gender, y=duration, fill=gender))+ 
  geom_violin(scale = "area")

```



## 2.3. Discrete X, Discrete Y

```{r}

df<- tibble(
  use = factor(c("wd","we","wd","we"), levels=c("wd","we")),
  duration = c(100,180,90,200),
  gender = factor(c("M","M","K","K"))
  
  )

df

```



### geom_count()

```{r}

ggplot(df, aes(use,gender)) +
  geom_count()

```



# 3. Graph Split

```{r}

df<- tibble(
  gender = factor(rep(c("M","F"), each=600)),
  mobile = c(rnorm(600, mean = 80), rnorm(600, mean = 82)),
  web = c(rnorm(600, mean = 82), rnorm(600, mean = 81)),
  job = factor(rep(c("A","B","C"), each = 400))
)

df

t<- ggplot(df, aes(mobile, web)) + geom_point()

t

# Let's split data by gender data vertically

t + facet_grid(. ~ gender)


# Let's split data by gender data horizontally

t + facet_grid(gender ~ .)


# Let's split data by gender and job data

t + facet_grid(job ~ gender) +
  
  geom_smooth(method = lm)

```



# 4. Time Series Visualization

```{r}

head(economics)


d<- economics

ggplot(d, aes(date, pop)) +
  geom_line()


# How to get subset of time date and visualize

s<- d %>% filter(date > as.Date("2010-1-1"))

ggplot(s, aes(date, pop)) +
  geom_line()


# How to add a dimension to time series. (Let's display unemployement as a dimension in timeline )
ggplot(d, aes(date, pop)) +
  geom_line(aes(size = unemploy/pop))


```



# 5. Multivariate Time Series Visualization

```{r}

library(tidyr)

df<- economics %>%
  select(date, psavert, uempmed) %>%
  gather(key = "variable", value = "value", -date)


head(df,3)

#------ geom_line()

ggplot(df, aes(date, value)) +
  geom_line(aes(color = variable), size =1)


#------geom_area()

ggplot(df, aes(date, value)) +
  geom_area(aes(color = variable, fill = variable),
            alpha = 0.5,
            position = position_dodge(.8),
            size =1)



  
```



# 6. Data Mining with Data Visualization

Below we will use Diamonds dataset to make data analysis and explore features in dataset.
```{r}

# Let's understand data

d<-diamonds


summary(d)


str(d)

# We see that price can be the most important variable that we wil benefit from

```


```{r}
# Let's see the price distribution of products


ggplot(d, aes(price)) +
  geom_histogram(bidwth = .5)


# Let's see the avg price by color and clarity separately

d %>%
  group_by(color) %>%
  summarise(n = n(), mean(price))


d %>%
  group_by(clarity) %>%
  summarise(n = n(), mean(price))



```


Let's continue to explore by color
```{r}
# Let's visualize color-price correlation
# (Note: it is discrete data)

k<- d %>%
      group_by(color) %>%
      summarise(n = n(), avg_price = mean(price))


ggplot(k, aes(color, avg_price)) +
  geom_bar(stat = "identity",
           color = "black",
           fill = "orange")

```


```{r}
# Let's see the price distribution by color distribution

ggplot(d, aes(price, fill =color)) +
  geom_density(alpha = .5)


str(d) # let's go over related data with the visualization together to interpret.



```

