---
title: "DataFrame"
author: "Vedat Ozkan"
date: "December 4, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# 1. What is DataFrame?


# 2. DataFrame Formating

## Create DataFrame

```{r}

df<- data.frame(
  
  store_id = c (1:5),
  region = c("Mecidiyekoy","besikTas","kadikoy","Bagcilar","Gop"),
  avg_profit = c(60,50,61,72,89),
  stringsAsFactors = FALSE
  
  )

df

str(df)

summary(df)

nrow(df)

ncol(df)

dim(df)

```

## Combine DataFrame 

```{r}

df2<- data.frame(
  
  store_id = c(1:5),
  past = c(2,4,5,3,6),
  emp_count = c(23,43,21,11,15)
  
)

# joins
merge(df,df2,by = "store_id")


cbind(df, df2)



```


## Add / Delete Variable

```{r df_all}

df_all<- merge(df, df2, by = "store_id")

df_all

df_all$avg_cust<- c(12,34,21,12,45)

df_all$product_name<- c("shoe","bag","skirt","tshirt","jacket")

df_all

## how to delete variable

df_all$city<- c("Arlington", "Fairfax", "Chantilly","DC","McLean")

df_all$city<- NULL

df_all

## how to do feature engineering

df_all$profit_by_emp<- df_all$avg_profit/df_all$emp_count

df_all

```

## Add / Delete Row

```{r}

df

rbind(df, c(6,"Beyoglu",99)) # add

df<- df[-6] # delete, or df<-df[1:5]

```


## Dataframe Naming

```{r}

colnames(df)<- c("n","s","k")

rownames(df)<- c("one","two","three","four","five")

df

# Now, let's make lower case 

df$s<- tolower(df$s)

#let's get only n and k columns / or remove s column

df<- df[,c("n","k")]

df

# How to assign columnnames to rownames


library(dplyr) # needed for %>% <==This sign called "Chain Operator"

library(tibble) # needed for remove_rownames

df_all<- df_all %>% remove_rownames %>% column_to_rownames (var="region")


# %>% kendinden onceki girdiyi kendinden sonraki cikti olarak atar

```


## Convert Lists / Matrices into Dataframe

```{r}

m<- matrix(1:9, 3,3)

df<-as.data.frame(m)

is.matrix(df)

is.data.frame(df)

#-------------

num<- c(1,2,3)
letter<- c("a","b","c")
log<- c(TRUE,FALSE,TRUE)

liste<- list(num, letter, log)

df1<-as.data.frame(liste)

```


## Convert Data Types in dataframe

```{r}

str(df_all)

df_all$product_name<- factor(df_all$product_name)

# note: if we need to get a variable as categorical we must convert it into factor

# Also let's define if the categorical data is ordinal or nominal

df_all$product_name<- factor(df_all$product_name, ordered = TRUE)

df_all$avg_cust <- as.integer(df_all$avg_cust) 


```


# 3.DataFrame Component Operations



## Dataframe Variable Selection
```{r}
library(ggplot2)

mpg

summary(mpg)

str(mpg)


mpg$model

attach(mpg) # you can directly tour in mpg without writing mpg$ each time, however, it can cause comfusing. Not recommended to use!

model

detach(mpg) # retrieves attach() function

# how to select more than one variable

data.frame(mpg$manufacturer, mpg$model)

head(data.frame(mpg$manufacturer, mpg$model)) # let's see first 6 rows.

```


## Dataframe Index and Element Selection

```{r}

mpg[1:10]

mpg[1,1]

mpg[1:10,1 ]

mpg[mpg$manufacturer == "audi" & mpg$cyl == "6" , ]

mpg[mpg$manufacturer == "audi" & mpg$cyl == "6" & mpg$year > 1999, c("model", "drv")]

```


## Dataframe Element Change

same operations with lists

```{r}



```


## Dataframe Sorting

```{r}

df_all[order(df_all$past, decreasing = TRUE),]



```


# 4. DataFrame Evaluation


```{r}

df_all$avg_profit*10 # we already learned it.


```


# 5. Apply Family

## apply()

```{r}

apply(matrix(1:9,3,3), 1, mean)


fap<- mpg[, c("displ","cyl","cty")]

fap

apply(fap, 1, sum)

apply(fap, 2, sum)
```


```{r}

sum(mpg$displ)

sum(mpg$cyl)

sum(fap[1,])

```


```{r}
# This is functional language, so we don't like use for-loop in R. We prefer apply() function

bos<- NULL

for (i in 1:nrow(fap)) {
  
  a<- sum(fap[i,])
  bos<- rbind(bos, a)
}


source("for.R")

```


```{r}
# R daki kod ve donguleri karsilastirmak ve test etmek icin asagidaki fonksiyonb kullanilir.

# ozellikle buyuk projelerde

microbenchmark(source("for.R"), apply(fap, 1, sum))

# YUKARIDAKI ISLEMI YAPMAMIZIN NEDENI DONGULERLE AYNI ISI YAPAN APPLY() FONKSIYONUNUN PERFORMANSINI KIYASLAMAK.
#APPLY() fonksiyonununayni isi yapan dongulerden cok daha yuksek performansa saahip oldugunu gorebiliriz

```

## lapply()

l letter stands for list
```{r}


l<- list( a = rnorm(90),
          b = rbeta(10,1,2),
          c = 1:5000
        )

lapply(l, mean)

```

## sapply()

s letter stands for simple. sapply() converts list into vectors

```{r}

l<- list( a = rnorm(90),
          b = rbeta(10,1,2),
          c = 1:5000
        )

sapply(l, mean)

```


## vapply()

```{r}

vapply(l, mean, double(1))

```


## tapply()

ensures for categorical variables to describe data / to factorize

```{r}

str(mpg)

tapply(mpg$cyl, mpg$class, mean)

```


## by()

ensures more detail in factorize

```{r}

by(mpg[,c("displ", "cty","hwy")], mpg$cyl, colMeans)

```



# 6.Map Family

```{r}

library(purrr)

mapp<- mpg[,c("displ", "cty","hwy")]

map_dbl(mapp, mean)

```


```{r}

library(dplyr)

mtcars %>%
  split(.$cyl) %>%
  map(~ lm(mpg ~ wt, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")
  

```



# 7. Modern Dataframe: Tibble



## First difference: Type style


```{r}

data.frame(
  
  a = 1:10000,
  b = 1:10000
)


tibble(
  
  a = 1:10000,
  b = 1:10000
)

iris # iris is a sample dataset

as_tibble(iris)
```


## Second difference: Subcluster Operations


























